---
title: "Create and Embed an Instant Flipbook"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, October 2020"
output: html_document
---

This template aimed at doing the following:  preparing a flipbook within an .Rmd file, which when rendered will *embed* the prepared flipbook in the resulting html.  Narrative might be interspursed between the code-movies.  (right now, only one code movie is working... got to figure out what's going wrong when we try for 2 imbedded flipbooks.)

How we get this done:

1. write some code in a code chunk
2. save that code out as text
3. create an "instant flipbook"
    - create an .Rmd file that will bring in code
    - render that file to an html xaringan slide show
4. embed the flipbook upon knitting the container doc to .html


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
```

### 1. Write code

Write some code in a code chunk to be code-movied. You can set echo to false later on.  This code chunk is named "my_code"

```{r my_code, eval = F}
cars %>% 
  ggplot() +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(size = 7) +
  aes(color = dist)
```

### 2. Save out code

Refer to code chunk to get code and save it out.  (Here I'm calling the code text temp.R, but maybe this should be a real temp file?).

The code will be pulled into the instant1.Rmd file and rendered to a xaringan slideshow w/ step-by-step reveal.  

```{r}
knitr::knit_code$get("my_code") %>% 
  paste(collapse = "\n") %>% 
  writeLines("temp.R")
```

### 3. Prepare instant flipbook


```{r}
write_instant_flipbook(file = "my_instant1.Rmd")
rmarkdown::render("my_instant1.Rmd", quiet = T)
```

### 4. Embed instant flipbook


The html is embedded:

```{r}
knitr::include_url(url = "my_instant1.html", height = 360)
```


### 4.b use xaringan extra...



```{r}
# another great option is to use the xaringanExtra package to embed
# install with remotes::install_github("gadenbuie/xaringanExtra")
xaringanExtra::embed_xaringan(
  url = "my_instant1.html",
  ratio = "16:10"
)
```



---


## Appendix

### What the .Rmd looks like for the instant flipbook

```{r}
knitr::include_url(url = "my_instant1.Rmd", height = 360)

```


### How we might bundle up steps 2 and 3

Some of the above could be contained in a single function `create_instant_flipbook`.  




```{r, eval = F}
create_instant_flipbook <- function(chunk_name, flipbook_name){
  
  # prepare code
  knitr::knit_code$get(chunk_name) %>% 
  paste(collapse = "\n") %>% 
  writeLines("temp.R")
  
  # pr
  flipbook_name_rmd <- paste0(flipbook_name, ".Rmd")
  flipbook_name_html <- paste0(flipbook_name, ".html")
  
  write_instant_flipbook(file = flipbook_name_rmd)
  rmarkdown::render(flipbook_name_rmd, quiet = T)
  
}

```

## What else

- It could be cool to allow a bunch of xaringan extra settings to be passed to the instant flipbook in the create_instant_flipbook function - to allow more customization.  

- you might allow `xaringanExtra::use_share_again()` to be turn on this function (the write_instant_flipbook function would have to change.)


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
