---
title: "Cool with statistical independence"
subtitle: "Basic bivariate statistical tests (and false positives)"
author: "Gina Reynolds"
date: "September/October, 2019"
output:
  # html_document:
  #   toc: no
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["xaringan-themer-special.css", "ninjutsu"]
    nature:
      beforeInit: "https://platform.twitter.com/widgets.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include = FALSE, eval = T}
knitr::opts_chunk$set(comment = "", warning = F, message = F, cache = T)
library(flipbookr)
```



---
# Statistical independence

> In probability theory, two events are independent, statistically independent, or stochastically independent if the occurrence of one does not affect the probability of occurrence of the other.

> Similarly, two random variables are independent if the realization of one does not affect the probability distribution of the other.

---
## "A statistically significant relationship"

- When you have *evidence* of statistical *dependence*.  The probability of observing of observing such a strong relationship --- or stronger --- under *statistical independence* (often the null) is quite small.  

https://www.youtube.com/watch?v=gSyGVDMcg-U

---
## Hypotheses

Hypothesis testing:

- A null hypothsis: Ho (Often statistical *independence*)
- Alternate hypothesis: Ha (Statistical *dependence*)

---
## Rejecting the null hypothesis

- Comparison will be made in relation to what we would expect to observe *under the null* (no relationship).  Such a strong relationship is unlikely to be observed by chance in a sample, if there is *not* dependence.  

---

- If data *deviates* from expecations about null

   - we reject the null (there is no relationship - statistical independence)
   - and accept the alternative (there is a relationship - statistical dependence)

---
## P-Value

The probability that a relationship as strong or stronger would have been observed under the null.

http://fivethirtyeight.com/features/not-even-scientists-can-easily-explain-p-values/ - but they all can...

---
## Fisher's exact test

```{r, eval = F}
number_correct = rep(NA, 1000000)
for (i in 1:1000000){
  guess <- sample(c(rep("tea first", 4), rep("milk first", 4)))
  truth <- sample(c(rep("tea first", 4), rep("milk first", 4)))
  number_correct[i] = sum(guess == truth)
}

table(number_correct)

table(number_correct)/10000
```






---

# Two statistically *independent* continuous variables



```{r continuous0, echo = F, eval = F, out.width="20%", comment=""}
library(tidyverse) 
set.seed(98735)
tibble(tail_length = rnorm(n = 50), 
       brain_weight = rnorm(n = 50)) -> 
my_data; my_data  #BREAK

ggplot(data = my_data) +
  aes(x = tail_length) +
  aes(y = brain_weight) +
  geom_point(color = "steelblue",
             alpha = .8) #BREAK
cor.test(x = my_data$tail_length, 
              y = my_data$brain_weight) #BREAK
```

---

`r chunk_reveal("continuous0", break_type = "user")`



---


```{r, echo = F}
# partial knit chunks

reveal_sequence <- function(chunk_name, num_ensembles = 10, seed = sample(1000:2000, num_ensembles, replace = F)) {
  # Create slide for lines 1:N for each line N in the given chunk
  # break_points <- seq_along(knitr:::knit_code$get(chunk_name)) # original code, breaks for each line

    out <- glue::glue(
      "class: split-40",
      "count: false",
      "```{r set_{{chunk_name}}_{{seed}}_1, echo = F}",
      "set.seed({{seed}})",
      "```",
      ".column[.content[",
      "```{r code_{{chunk_name}}_{{seed}}, eval=FALSE, code=knitr::knit_code$get('{{chunk_name}}')}",
      "```",
      "]]",
      # "```{r set_{{chunk_name}}_{{seed}}_2, echo = F}",
      # "set.seed({{seed}})",
      # "```",
      ".column[.content[",
      "```{r output_{{chunk_name}}_{{seed}}, echo=FALSE, code=knitr::knit_code$get('{{chunk_name}}'), , comment='   '}",
      "```",
      "]]",
      .open = "{{", .close = "}}", .sep = "\n"
    )

  glue::glue_collapse(x = out, sep = "\n---\n")
}

apply_repeat_reveal <- function(...){
  paste(knitr::knit(text = reveal_sequence(...)), collapse = "\n")
}



```


```{r continuous, echo = F, eval = F, out.width="20%", comment=""}
library(tidyverse)
tibble(tail_length = rnorm(n = 50), 
       brain_weight = rnorm(n = 50)) ->
my_data; my_data
#
#
#
#
#
#
#
#
#
#
#
#
#
ggplot(data = my_data) +
  aes(x = tail_length) +
  aes(y = brain_weight) +
  geom_point(color = "steelblue",
             alpha = .8) +
  theme(panel.background = 
          element_rect(color = 
                         ifelse(cor.test(my_data$tail_length, my_data$brain_weight)[[3]]<.05, "red", "black"),
                       size =
                         ifelse(cor.test(my_data$tail_length, my_data$brain_weight)[[3]]<.05, 8, 3)))
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
cor.test(x = my_data$tail_length, 
              y = my_data$brain_weight)
```

---

`r chunk_reveal("continuous", break_type = 20)`



---

```{r c_and_d2, echo = F, eval = F}
tibble(sex = sample(x = c("male","female"), 
                  size = 50, 
                  replace = T)) %>% 
  mutate(height = rnorm(n = 50, sd = .5) + 5.5) ->
height_data; height_data
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
# visualization
ggplot(height_data) +
  aes(x = sex) +
  aes(y = height) +
  aes(group = sex) +
  geom_boxplot() +
  geom_jitter(height = 0,
              width = .02) +
  stat_summary(fun.y = mean,
               geom = "point",
               col = "goldenrod3",
               size = 5) 
#
#
#
#
#
#
#
#
# statistical test
t.test(height_data$height ~ # the continuous variable by (~)
         height_data$sex) # the discrete variable







```

---




---

`r chunk_reveal(chunk_name = "c_and_d2", break_type = 20)`

---


# ANOVA (analysis of Variance)


---

```{r c_and_d3, echo = F, eval = F}
tibble(sex = sample(x = c("1st","2nd","3rd","Crew"), 
                  size = 50, 
                  replace = T)) %>% 
  mutate(height = rnorm(n = 50)) ->
height_data; height_data
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
ggplot(height_data) +
  aes(x = sex) +
  aes(y = height) +
  aes(group = sex) +
  geom_boxplot() +
  geom_jitter(height = 0,
              width = .02) +
  stat_summary(fun.y = mean,
               geom = "point",
               col = "goldenrod3",
               size = 5) 
#
#
#
#
#
#
#
#
#
#
#
#
# statistical test
TukeyHSD(aov(height_data$height ~ # the continuous variable by (~)
         height_data$sex)) # the discrete variable







```


---

`r chunk_reveal(chunk_name = "c_and_d3", break_type = 20)`





---


# Chi-squared test 

$$\chi$$

Two-by-two

```{r d_and_d, echo = F, eval = F}
library(tidyverse)
tibble(sex = sample(x = c("male","female"), 
                  size = 100, 
                  replace = T), 
       handed = sample(x = c("left", "right"), 
                  size = 100, 
                  replace = T, 
                  prob = c(.3,.7))) ->
my_data ; my_data
#
#
#
#
#
#
#
#
#
#
library(ggmosaic) # geom_mosaic() and product()
ggplot(data = my_data) +
  geom_mosaic(aes(x = product(sex),
                  fill = handed))
#
#
#
#
#
#
#
#
#
# contingency table or two by two table
table(x = my_data$sex, y = my_data$handed)
#
#
#
#
#
chisq.test(x = my_data$sex, 
           y = my_data$handed) 
```

---



`r chunk_reveal(chunk_name = "d_and_d", break_type = 20)`



---

```{r lady_tea, echo = F, eval = F}
library(tidyverse)
tibble(truth = sample(rep(c("tea", "milk"), 4))) %>% 
  mutate(guess = sample(rep(c("tea", "milk"), 4))) ->
tea_data ; tea_data
#
#
#
#
#
#
#
#
#
#
#
#
#
#
library(ggmosaic) # geom_mosaic() and product()
ggplot(data = tea_data) +
  geom_mosaic(aes(x = product(truth),
                  fill = guess))
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
#
chisq.test(x = tea_data$truth, 
           y = tea_data$guess) 
```


---

---



`r chunk_reveal(chunk_name = "lady_tea", break_type = 20)`

---






---

```{r d4_d2, echo = F, eval = F}
library(tidyverse)
tibble(class = sample(x = c("1st","2nd","3rd","Crew"), 
                  size = 50, 
                  replace = T)) %>% 
  mutate(survive = sample(x = c("Survive", "Perish"), 
                          size = 50, 
                          replace = T, 
                          prob = c(.3, .7))) ->
fake_titanic ; fake_titanic
#
#
#
#
#
#
#
#
#
#
library(ggmosaic) # geom_mosaic() and product()
ggplot(data = fake_titanic) +
  geom_mosaic(aes(x = product(class),
                  fill = survive)) -> g1
ggplot(data = fake_titanic) +
  aes(x = class) +
  aes(y = survive) +
  geom_count(alpha = .5, color = "blue") +
  geom_jitter(width = .2, height = .2) -> g2
library(patchwork)
g1 + g2
#
#
#
#
#
#
#
#
#
#
#
#
#
chisq.test(x = fake_titanic$class, 
           y = fake_titanic$survive) 
```

---

`r chunk_reveal(chunk_name = "d4_d2", break_type = 20)`




```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 30%}
```